BOLD Parameter Estimation Using Particle Filters

### 2010-08-11 doc/thesis/conclusion.tex
more thesis gah

\chapter{Conclusion}
\label{sec:Conclusion}
This work has proposed and demonstrated the use of a particle filter
for the estimation of BOLD model parameters. Because many of the parameters
were under constrained from the point of view of the BOLD response, a
full posterior estimate was a more logical approach to parameter estimation
than a single point estimate of the parameters. The particle filter also converged
quickly to correct output estimates, and thus provides a framework
for future real-time FMRI experiments. Final BOLD estimates were also comparable
to the results of SPM.

>>>>

\chapter{Conclusion}
\label{sec:Conclusion}
This work has demonstrated the use of the particle filter to
learn parameters of the BOLD model. Since the inception of the
BOLD equations, many attempts have been made to learn the parameters,
given FMRI data. These attempts have typically either been extremely 
slow or made extensive approximations. While the particle filter method
is not quick, 60 seconds to analyze each voxel is well within the capabilities
of the typical research lab.

One significant finding of this work, that would not be clear without 
calculating a full posterior distribution, is the interplay 
between the parameters. The results of simulations clearly demonstrate
that identifying a single set of parameters is not possible with this 
model. Although sensitivity tests in \cite{Deneux2006} certainly hinted
at this, the current work clearly demonstrates this fact. Therefore,
any single estimate of parameters is insufficient for analysis. As such,
approaches to the BOLD model that do not treat the parameters as distributions
will not be able to overcome the inherent indeterminability of the
parameters. Besides the Unscented Kalman Filter, this is the only approach
that accomplishes this task without recalculating the parameters over and 
over to build a distribution. The Unscented Kalman Filter, though, is limited
to Gaussian estimation.

### 2010-08-11 doc/thesis/discussion.tex
more thesis gah

Although point estimates of parameters are not dependable estimates of the 
true parameters, they still useful. Analysis of differences in parameters
could be medically relevant for instance. Additionally, a single estimate for parameters is still able
to form an accurate estimate of the BOLD signal. This trait bridges the gap with earlier types 
of activation tests, such as the statistical
parametric mapping. As the results in \autoref{sec:RealData} show, the heatmaps, especially
those of mutual information, closely resembled the results of SPM. While the activation
tests were more sensitive, they were also more prone to false positive. However, it is
worth noting that certain enhancements could be useful in reducing false positives; for instance
by using the maximum likelihood, or median of the final distribution.  Future work may 
shed further light on these techniques.

>>>>

Although point estimates of parameters are not dependable estimates of the 
true parameters, they still useful. Analysis of differences in parameters
could be medically relevant for instance. Additionally, a single estimate for parameters is still able
to form an accurate estimate of the BOLD signal. This trait bridges the gap with earlier types 
of activation tests, such as the statistical
parametric mapping. As the results in \autoref{sec:RealData} show, the heatmaps, especially
those of mutual information, closely resembled the results of SPM. While the activation
tests were more sensitive, there were some false positive, though at this point it is
difficult to quantify. It is
worth noting that certain enhancements could be useful in reducing false positives; for instance
by using the maximum likelihood, or median of the final distribution.  Future work may 
shed further light on these techniques.

### 2010-08-11 doc/thesis/resultsB.tex
more thesis gah

>>>>

\section{FMRI Configuration}
For the FMRI data discussed in \autoref{sec:RealData}, tests were 
performed on a right handed volunteer using a Gradient EPI sequence
on a 1.5T GE scanner with a 2.1s TR. Slices were 5mm thick and echo
time was (TODO). The subject was then presented with either a single or
double flash, at randomly dispersed time points over the course of
5 minutes, 10 seconds. The FMRI began 18.9 seconds
before the beginning of the experiment, to allow for transients 
in to image to settle out. The timing of the flashes are shown in 
\autoref{tab:flashes}, with time 0 corresponding to the beginning of
the 10th TR (so 9 images were dropped).

\begin{table}[t]
\centering
\begin{tabular}{|c || c |}
\hline 
Single Flash Times & Double Flash Times \\
\hline 
.7 & 2\\
2.7 & 21\\
5.7 & 28\\
12.7 & 52\\
\hline 
\end{tabular}
\caption{Timing of single and double flashes.}
\label{tab:flashes} 
\end{table}

After dropping the first 9 volume, each of the remaining volumes was
co-registered with the first volume in time. At this point the SPM method
diverges from the experimental method. For the particle filter, detrending 
was then applied as discussed in \autoref{sec:Detrend} and the resulting
data was processed with the particle filter. To generate the SPM output
in this section, the co-registered data was passed through a high pass
filter with a cut off of (TODO). The data was then smoothed spatially 
with a Gaussian kernel (FWHM 8x8x8 voxels). A Statistical Parametric Map
was then generated using SPM8. All settings used in SPM8 were the defaults
for analyzing FMRI data.
